A Machine Learning Approach to MLB Catcher Framing

### 2021-12-10 Slides.Rmd
Added ridge regression and LASSO to slides

```{r setup, include=FALSE}
library(tidyverse)
library(gbm)
library(randomForest)
knitr::opts_chunk$set(echo = FALSE)
```

@@

log_reg_fit <- glm(strike ~ ., data=data_no_catchers[train, ], family='binomial')
summary(log_reg_fit)

pred <- predict.glm(log_reg_fit, type='response', newdata = data_no_catchers[test, ])
pred[pred >= 0.5] <- 'strike'
pred[pred < 0.5] <- 'ball'
sum(pred=='strike')/length(pred)

@@

conf_mx
error
```

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(glmnet)
library(gbm)
library(randomForest)
knitr::opts_chunk$set(echo = FALSE)
```

@@

log_reg_fit <- glm(strike ~ ., data=data_no_catchers[train, ], family='binomial')

pred <- predict.glm(log_reg_fit, type='response', newdata = data_no_catchers[test, ])
pred[pred >= 0.5] <- 'strike'
pred[pred < 0.5] <- 'ball'

@@

conf_mx
1-error
```

- The variables with significant results were:
- Pitch type, release speed, z release position
- Whether the batter is a lefty/righty
- Count
- Where the pitch landed
- Most of the innings
- Batter height
- Whether it was a home game

## Ridge Regression

```{r}
x <- model.matrix(strike ~ ., data = data_no_catchers)[, -1]
y <- data_no_catchers$strike

lambda <- seq(0.001, 10, length=100)

#Create training and test data set
train <- sample(1:nrow(x), nrow(x)/2)
test <- (-train)
ridge_reg_fit <-  glmnet(x[train,], y[train], alpha = 0, lambda = lambda, family='binomial')

#Choose lambda
cv_ridge <- cv.glmnet(x[train, ], y[train], alpha=0, family='binomial')
lambda_choice <- cv_ridge$lambda.min

#Get predictions
ridge_predict <- predict(ridge_reg_fit, s = lambda_choice, newx = x[test, ], type = 'class')

#Confusion matrix
ridge_conf_mx <- table(pred=ridge_predict, true=y[test])
ridge_conf_mx
error <- (ridge_conf_mx[2] + ridge_conf_mx[3])/length(ridge_predict)
1-error
```

## LASSO

```{r, cache=TRUE}
#LASSO
lasso_fit <- glmnet(x[train, ], y[train], alpha = 1, lambda = lambda, family = 'binomial')
cv_lasso <- cv.glmnet(x[train, ], y[train], alpha = 1, family='binomial')

#Lambda choice
lasso_lam <- cv_lasso$lambda.min

#Get predictions
lasso_predict <- predict(lasso_fit, s = lasso_lam, newx = x[test, ], type = 'class')

#Get coefficients
lasso_coeff <- predict(lasso_fit, s = lasso_lam, type = 'coefficients')

non_zero_lasso_coef <- lasso_coeff[lasso_coeff != 0]
non_zero_lasso_coef
#Confusion matrix
lasso_mx <- table(pred=lasso_predict, true=y[test])
lasso_mx
error <- (lasso_mx[2] + lasso_mx[3])/length(lasso_predict)
1-error
```

- Some of the non-zero coefficients include:
- Pitch type
- Count
- Outs and Inning
- Batter height
- Whether the game is home or away
- Where the pitch landed
